---
title: "ABC Supermarkets Survey"
author: "https://github.com/VictoriaUniversity-AU/assessment-1-Cathy-2022.git"
date: '2022-08-08'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(dplyr)
library(ggplot2)
library(widyr)
#install.packages("lubridate")
library(lubridate)
library(tidytext)
library(tidymodels)

```
### Use parse_number() to covert purchase.price, total.cost,sold.price and total.sales to number, remove $ signs
```{r}
ABC_Supermarket <- read.csv("Survey data.csv")
ABC_Supermarket %>% 
  group_by(suplier) %>% 
  ungroup()
  
ABC_df <- ABC_Supermarket %>% 
  mutate(purchase_price_num = parse_number(purchase.price),
         total_cost_number = parse_number(total.cost),
         sold_price_num = parse_number(sold.price),
         total_sales_num = parse_number(total.sales)
         ) 

```
### 1.What are the top 5 company items that have more demand based on items sold

```{r}
top_items <- ABC_Supermarket %>% 
  group_by(suplier,Itemname) %>% 
  summarise(
    number=n(),
    item_sum = sum(Item.sold, na.rm = TRUE)
  ) %>% 
  ungroup()

top_items <- top_items %>% 
  arrange(-item_sum) %>%
  head(5) %>% 
  select(Itemname,suplier, item_sum)
top_items

top_items %>% 
  ggplot(
  aes(fct_reorder(Itemname, item_sum), item_sum, fill = suplier)) +
  geom_col(alpha = 0.8) 

```
### What are the top 5 company sales that have more demand based on items sold
```{r}
top_sales <- ABC_df %>% 
  group_by(suplier,Itemname) %>% 
  summarise(
    number=n(),
    total_sales_sum = sum(total_sales_num, na.rm = TRUE),
    total_cost_sum = sum(total_cost_number, na.rm = TRUE)
  ) %>% 
  ungroup()

top_sales <- top_sales %>% 
  arrange(-total_sales_sum) %>%
  head(5) %>% 
  select(Itemname,suplier, total_sales_sum)
top_sales

top_sales %>% 
  ggplot(
  aes(fct_reorder(Itemname, total_sales_sum), total_sales_sum, fill = suplier)) +
  geom_col(alpha = 0.8) 
```
### 2.Finding out the top 5 profitable items and giving these items sales line chart to find it sale’s trend.
```{r}
ABC_df <- ABC_df %>%
  mutate(profit = total_sales_num - total_cost_number,
         year_month= Year*100 +Month.Number)

profit <- ABC_df %>% 
  group_by(Itemname) %>% 
  summarise(item_profit= sum(profit),
            item_sold= sum(Item.sold),
            ) %>% 
  ungroup()

profit %>% arrange(-item_profit) %>%
  head(5) 


top_profit <- ABC_df %>% group_by(Itemname,year_month) %>% 
  summarise(total_profit= sum(profit),
            total_sales = sum(total_sales_num)) %>% 
  filter( Itemname=="Squash" |
           Itemname == "Bananas" |
           Itemname == "Oranges" | 
           Itemname == "Lettuce" |
           Itemname == "Tomatoes" 
         )

top_profit %>% ggplot(
  aes(year_month,total_sales,colour=Itemname,fill=Itemname)
  )+
  geom_line()+
  facet_wrap(~ Itemname, scales = "free") +
  ggtitle("Top 5 Items Profitable Items")

```
### 3.	Finding out 5 lowest profitable items, giving the 5 lowest items sales line chart to find it sale’s trend.
```{r}
profit %>% arrange(item_profit) %>%
  head(5) 

low_profit <- ABC_df %>% group_by(Itemname,year_month) %>% 
  summarise(total_profit= sum(profit),
            total_sales = sum(total_sales_num)) %>% 
  filter( Itemname=="Celery" |
           Itemname == "Mushrooms" |
           Itemname == "Cheese" | 
           Itemname == "Sour cream" |
           Itemname == "Cucumber" 
         )

low_profit %>% ggplot(
  aes(year_month,total_sales,colour=Itemname,fill=Itemname)
  )+
  geom_line()+
  facet_wrap(~ Itemname, scales = "free") +
  ggtitle("Lowest 5 Profitable Items")
```
### 4.Finding the total sales per month for each year, finding which month has the highest sales comparing to other months. 
```{r}

sale_pm <- ABC_df %>% 
  group_by(year_month) %>% 
  summarise(sales_per_month = sum(total_sales_num)
    ) %>% 
  arrange(year_month)

sale_pm %>% 
  ggplot(
  aes(year_month,sales_per_month,fill=sales_per_month)
  )+
  geom_line()+
  ggtitle("Total Sales per Month")

```
### Comparing total sales for each year.

```{r}
sales_year <- ABC_df %>% 
  group_by(Year) %>% 
  summarise(sales_per_year = sum(total_sales_num)
  ) 
sales_year %>%  ggplot(
  aes(x = Year, y = sales_per_year, fill = sales_per_year)  
  )+
  geom_col() +
  labs(y = "Total Sales")
  ggtitle("Total Sales per Year")
growth_rate <- sales_year %>% mutate(
         Diff_year = Year - lag(Year), 
         Diff_growth = sales_per_year - lag(sales_per_year), 
         Rate_percent = (Diff_growth / Diff_year)/sales_per_year * 100)
  
```
### 5.Using lm() function to find is there any relationship between feedback and total sales. Finding out the 5 items which have lowest total sales and finding out their sentiment.
```{r}
ABC_lm <- ABC_df %>% 
  filter(!is.na(total_sales_num))

lm_abc = lm(total_sales_num ~other.feedback, data=ABC_lm)
summary(lm_abc)

ABC_df %>% 
  group_by(Itemname,suplier) %>% 
  summarise(total_sales= sum(total_sales_num),
            item_sold= sum(Item.sold),
            ) %>% 
  ungroup() %>% arrange(total_sales) %>%
  head(5)
 
 low_total_sales <- ABC_df %>% 
   filter(grepl("Glendal Foods", suplier, ignore.case = TRUE) & Itemname =="Celery" |
          grepl("Mediterranean Wholesalers", suplier, ignore.case = TRUE) & Itemname =="Yogurt" |
          grepl("Bruno Fine Foods", suplier, ignore.case = TRUE) & Itemname =="Lettuce" |
          grepl("Glendal Foods", suplier, ignore.case = TRUE) & Itemname =="Yogurt" |
          grepl("Quality Foods", suplier, ignore.case = TRUE) & Itemname =="Cucumber"
          ) %>% 
   select(Itemname,suplier,total.sales,other.feedback)
 
feedback_token <- low_total_sales %>% 
  unnest_tokens(word,other.feedback)

nrc<- get_sentiments("nrc")
feedback_token <-feedback_token %>%
  inner_join(nrc) %>% 
  rename(sentiment_nrc=sentiment) %>% 
  select(Itemname,suplier,sentiment_nrc)
```
### 6.Which suppliers has the highest total sales?  Did they have positive feedback? 
```{r}
sales_suplier <- ABC_df %>% filter(!is.na(total.sales)) %>% 
  group_by(suplier) %>% 
  summarise(total_sale=sum(total_sales_num)) %>% 
  arrange(-total_sale) 

sales_suplier %>% ggplot(
  aes(fct_reorder(suplier, total_sale),total_sale, fill = suplier)) +
  geom_col(alpha = 0.8) 

best_suplier <- ABC_df %>% 
  filter(grepl("Basfoods Direct", suplier, ignore.case = TRUE) &
        !str_detect(suplier, "Basfoods Direct - Somerton Store"))
best_suplier %>% distinct(Itemname)

best_token <- best_suplier %>% 
  unnest_tokens(word,other.feedback)

bing<- get_sentiments("bing")
best_token <-best_token %>%
  inner_join(bing) %>% 
  rename(sentiment_bing=sentiment) %>% 
  select(suplier, sentiment_bing)

best_token <- best_token%>%
  count(sentiment_bing, sort=TRUE)%>% 
  mutate(percent=paste0(round(n/sum(n)*100,2),"%"))
counts<-best_token %>% count(sentiment_bing, sort = TRUE) 
ggplot(best_token, aes(x = "", y = percent, fill = sentiment_bing)) +
  geom_col() +
  geom_text(aes(label = percent),
            position = position_stack(vjust = 0.5)) +
  coord_polar(theta = "y")+
  ggtitle("Sentiment bing for best supplier")
```
### Which suppliers has the lowest total sales?  Did they have negative feedback?
```{r}
ABC_df %>% 
  filter(grepl("Delta Sales", suplier, ignore.case = TRUE)) %>% 
  select(Itemname,suplier,total.sales,Item.sold,other.feedback)
```
### 7.Which supplier provided items which has the highest profit based on year.
```{r}
hiprofit_supplier <- ABC_df %>% 
  group_by(suplier,Year) %>% 
  summarise(total_profit=sum(profit)) %>% 
  arrange(total_profit)
hiprofit_supplier %>% ggplot(
  aes(total_profit, fct_reorder(suplier, total_profit), fill = Year)) +
  geom_col(alpha = 0.8) 

```
###8.Which item's profit increased highly in these years? Find out the top 5. Which item's profit decreased in these years.
```{r}
hiprofit_item <- ABC_df %>% 
  group_by(Itemname,Year) %>% 
  summarise(total_profit=sum(profit)) %>% 
  arrange(total_profit)

growth_profit <-hiprofit_item %>% 
  group_by(Itemname) %>% 
  mutate(
         Diff_growth = total_profit - lag(total_profit), 
         Rate_profit = (Diff_growth / lag(total_profit) /100)
  )%>% 
  arrange(-Rate_profit)  
  
growth_profit %>% ggplot(
  aes(Rate_profit, fct_reorder(Itemname, Rate_profit), fill = Itemname)) +
  geom_col(alpha = 0.8) 


```



