---
title: "Business Analysis"
author: "https://github.com/VictoriaUniversity-AU/assessment-1-Cathy-2022.git"
date: '2022-08-26'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(dplyr)
library(ggplot2)
library(widyr)
library(readxl)
```
### Import data
```{r}
df <- read_excel("Data file.xlsx")
str(df)
```

### 1.Identify items sold in the year 2017 Discuss what you have learned from the data outcome 
```{r}
df <- df %>% mutate(
  year = as.numeric(substr(`Order Date`, 1,4)),
  month = as.numeric(substr(`Order Date`, 6,7)),
  date = as.numeric(substr(`Order Date`,9,10))
)

df %>% count(year, sort = TRUE)

df_2017 <- df %>% filter(year==2017) %>% 
  count(typeCategory) %>% 
  arrange(-n) 

p <- df_2017 %>% filter(n>10) %>% 
  ggplot( 
  mapping = aes(
  x = typeCategory, fill = typeCategory, y = n))
p + geom_col()+labs(title="Type Category sold in 2017")
```
### 2. Does product price varies city to city? Identify data, analyze and discuss result.
```{r}
df %>% count(typeCategory, sort = TRUE) 
df %>% count(`Order City`)

df_popular <- df %>% group_by(typeCategory) %>% 
  filter(n()>40)

price_compare <- df_popular %>% group_by(`Order City`,typeCategory) %>% 
  summarise(
    count=n(),
    mean_price=mean(`Product Price`,na.rm=TRUE),
    median_price=median(`Product Price`,na.rm=TRUE),
    max_price=max(`Product Price`,na.rm=TRUE),
    min_price=min(`Product Price`,na.rm=TRUE)
  ) %>% arrange(typeCategory)

price_compare %>% filter(
  grepl("beleza_saude", typeCategory, ignore.case = TRUE)
  ) %>% 
  ggplot(
  aes(mean_price, fct_reorder(`Order City`, mean_price), fill = `Order City`)) +
  geom_col(alpha = 0.8) 

price_compare %>% filter(
  grepl("cama_mesa_banho", typeCategory, ignore.case = TRUE)
  ) %>% 
  ggplot(
  aes(median_price, fct_reorder(`Order City`, median_price), fill = `Order City`)) +
  geom_col(alpha = 0.8) 
```
### 3. Does type-Category vary city to city? Identify data, analyze and discuss results
```{r}
df %>% count(typeCategory)
df_city_type <- df %>% count(`Order City`, typeCategory)
df_city_type %>% count(`Order City`) %>% 
  arrange(-n)
```
### 4. Identify order where the customer use “credit_card” payment method? Identify data, analyze and discuss result 
```{r}
df_pay_credit <- df %>% count(`Order City`, `Payment Method`) %>% 
  filter(`Payment Method`=="credit_card") %>% 
  arrange(-n) %>% 
  rename(credit_n=n)

df_pay <- df %>% count(`Order City`)
df_pay <- df_pay %>%    
  left_join(df_pay_credit) %>% 
   mutate(percent=round(credit_n/n,2)) %>% 
         arrange(-percent) 
df_pay %>% ggplot(
  aes(percent, fct_reorder(`Order City`, percent), fill = `Order City`)) +
  geom_col(alpha = 0.8) 
```
### 5. identify items’ product prices between 45 and 552. 
```{r}
df_5 <- df %>% filter(`Product Price`>45 & `Product Price`<552)

df_5 %>% count(`Order City`)%>% 
  arrange(-n)
df_5 <- df_5 %>% count(`Order City`,year) 

df_5 %>% ggplot(
  aes(x=n, y=`Order City`,color=year, fill=year), alpha=.6)+
  geom_col()+
  labs(title="Item sold for each city")
```

